Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid and Debt, Xinhua and Diplomatic Visits

### 2022-11-03 04_paper/1103_draft.Rmd
changes upfront

---
title: "Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits"
author: ["Lucie Lu","Miles D. Williams"]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2: 
    toc: false
classoption: fleqn
fontfamily: mathpazo
fontsize: 12pt
bibliography: ref_xinhua.bib
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \setlength{\mathindent}{15pt}
  - \setlength\parindent{24pt}
  - \setlength{\parskip}{0.0pt plus 1.0pt}
  - \doublespacing
  - \newtheorem{theorem}{Theorem}[section]
  - \newtheorem{lemma}[theorem]{Lemma}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{corollary}[theorem]{Corollary}
  - \newenvironment{proof}[1][Proof]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries Proof}]}{\end{trivlist}}
  - \newenvironment{definition}[1][Definition]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{example}[1][Example]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{remark}[1][Remark]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{!t}
  - \usepackage{caption}\captionsetup[table]{skip=5pt}
  - \usepackage{dcolumn}
  - \usepackage{rotating}
  - \usepackage{caption}\DeclareCaptionLabelFormat{AppendixTables}{A.#2}

abstract: "Absent formal avenues of transparency, how do China’s foreign aid and debt allocation relate to public diplomacy and international recognition efforts? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. We expect greater coverage of a developing country in a Chinese media outlet directed to aid giving, and greater bilateral diplomatic activity to correlate with greater debt allocation. To test this expectation, we leverage (1) AidData’s Chinese development finance datasets and (2) the meta-data of millions of Xinhua news articles in the Cline Center for Advanced Social Research. We find that China pays more media attention to aid recipients but engages more in diplomatic visits with debtors. Aid-recipient countries disproportionately get more media coverage on the state of their economy and developmental needs. Also, China is more likely to provide loans to diplomatically aligned countries. Our results dispel the myths of China as a ’rouge donor’ and Chinese ’debt-trap diplomacy.’ We suggest that much like traditional donors, China’s foreign aid presents a versatile foreign policy tool in international politics. "
---

@@

However, the goals of facilitating cooperation and winning international recognition as a rising power are difficult to accomplish in secret. This, therefore, raises a natural question. *How do China's foreign aid and debt allocation relate to public diplomacy and international recognition efforts?*

We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. Specifically, we consider outward-facing state-sanctioned news coverage and bilateral diplomatic missions from Beijing. Past research suggests that these two activities play special roles in complementing the foreign policy objectives that drive China's giving and lending.

Media coverage is one way that countries can generate attention and promote a positive spin on their activities to foreign audiences. As Nye [-@nye_public_2008-2] notes, daily communications that reach both domestic and foreign populations are a way to improve a country's soft power. Perhaps in recognition of this, Beijing has expanded its outward-facing state-sponsored media activities and news reporting over the past two decades. This has been done in part through the English version of *Xinhua*, China's top state-sponsored news agency.

In addition to expanding the reach and scope of its externally facing news media, China engages in more conventional forms of bilateral diplomacy via government-to-government diplomatic missions. Efforts to document Beijing's public diplomacy have revealed an expansive set of activities in which Beijing engages. These range from the financial to the cultural, and from elite-to-elite meetings to informational exchanges.

@@

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to the type of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.[^1] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

@@

While the findings overall are consistent with our hypotheses, they do so in ways that also problematize the neat distinction between China's use of aid and debt. In particular, we find that *Xinhua* coverage is disproportionately concentrated among China's ODA recipients, while no significant relationship exists between aid and diplomatic visits. Conversely, we find that debt is significantly linked to a greater frequency of bilateral diplomatic missions from Beijing, but no significant difference in *Xinhua* coverage. Together these findings reveal an intriguing pattern: Beijing likes to talk about its aid recipients but prioritizes diplomacy with its debtors.

Aid recipient countries disproportionately get more media attention from *Xinhua* via stories that report both on the state of their economy and their development needs. Notably, *Xinhua* does not just promote China's efforts to offer aid to those countries in need but also gives a rounded picture of those countries' economic conditions justifying their legitimate need for aid.

Meanwhile, the finding that loans correspond with diplomatic missions has special relevance in light of Beijing's Belt and Road Initiative (BRI). A 2018 BRI foreign policy assessment report from Washington [@hurley_examining_2019] highlights that lending indicates Beijing's economic interests and strategic dominance---China tends to lend to vulnerable states with debt distress in exchange for unfair deals and strategic concessions. Brautigam [-@brautigam_critical_2020] calls the Chinese "debt-trap diplomacy" a meme based on collective fear of the Chinese threat rather than sound empirical evidence. Our finding suggests that China is more likely to provide loans to diplomatically aligned countries. For instance, Russia, one of China's closest allies, receives the highest amount of lending on average among 142 countries but no aid at all between 2002 and 2017. In contrast, Iraq shows the opposite pattern: it receives the highest amount of aid from China during the period of our investigation, but it does not receive any loans. Concern about Chinese debt nonetheless raises a natural question related to the function of diplomacy in debtor countries. Does debt serve as a tool of diplomacy, or does Beijing wield diplomacy to cover a multitude of "debt-trap" sins? More research is needed to answer this question.

@@

Absent transparency in reporting, there are other tools at Beijing's disposal that may complement the diplomatic objectives behind its foreign aid. In this study, we explore the connections between China's bilateral development finance and other, *visible*, activities linked to diplomatic and legitimacy-seeking objectives. We focus in particular on two factors: (1) diplomatic visits from China to developing countries and (2) coverage of developing countries in external-facing media (that is, media coverage intended for foreign, rather than domestic, audiences).

We summarize these factors in greater detail in the following sections, but before considering them, it is necessary to first summarize the current landscape of China's overseas financing. In particular, it is important to draw a distinction between Beijing's ODA-like expenditures and its other official flows (OOF). The former we may call aid (or development assistance), and the latter we may call debt (or loans). As @dreher_banking_2022 note in their groundbreaking and comprehensive book on Chinese aid and loans, Beijing uses aid and debt to accomplish different goals, and it further does not finance them in equal proportion, both in terms of total spending and in terms of how it targets funds across developing countries. @dreher_banking_2022 argue that aid tends to be most associated with China's foreign policy objectives, while debt tends to correspond to its market interests. What do these differences imply for Beijing's bilateral diplomacy and state-sponsored foreign media?

On the one hand, since aid is supposed to have more direct links to China's foreign policy goals, it follows that China will complement its aid giving with visible acts of diplomacy and efforts to increase the salience of aid recipients in foreign audiences. On the other hand, while aid and debt may have distinct goals, such a tidy distinction may be too simplistic. Perhaps more importantly, some of the market oriented objectives that drive China's global debt might benefit from public displays of diplomacy. On this front, recent research suggests that Beijing uses its debt to ensure access to natural resources that China lacks domestically. As noted above, @dreherEtAl2021 find that Beijing targets more of its loans in resource rich developing countries, a strategy that @dreher_banking_2022 contend is meant to guarantee access to said resources. When Beijing's policy banks provide these loans, developing country governments collateralize their oil or rare minerals, all but assuring access to these resources for Beijing since many of these countries are unable to directly repay the loans[@dreher_banking_2022].

What this implies in the aggregate is that aid from China tends to be most strongly determined by both development and political factors, while debt is most strongly determined by market-based factors. For the two outcomes discussed in the following sections, this suggests that Chinese ODA-like expenditures more so than OOF should be strong predictors of other complementary and visible foreign policy activities. Meanwhile, given the more market-oriented (and by extension more controversial) application of debt, we should expect a much weaker association to exist between OOF and the factors summarized below. While visibility may be a virtue for aid, the opposite may be true for debt. The next section summarizes these outcomes and their relevance as measures of foreign policy-related activities in the context of China's overseas financing.

@@

We have made the case that both foreign state-sponsored media and diplomatic visits can be used to sway foreign public opinion in favor of Beijing. By covering developing countries that receive Chinese development financing, *Xinhua* editors can paint a positive picture of Beijing's financing and, in so doing, take steps to promote China's international image. In addition, by sending regular diplomatic missions to developing countries, Beijing is able to strengthen its bilateral relationships with countries in the Global South helping to cement the close bilateral ties the offer of development assistance is, in part, indented to promote.

@@

# include_graphics(
#   here('06_figures/coverages_combined.jpg'))

@@

# include_graphics(
#   here('06_figures/visits_combined.jpg'))

@@

To put these estimates in more substantive terms, it will be helpful to show some computed marginal effects. Table \@ref(tab:mareff) shows the predicted change in each of the outcomes given a percent increase in either aid or debt in Pakistan in 2017. Cell entries in the first column denote the \% change in the number of times Pakistan appears in *Xinhua* in 2017 while entries in the second column denote the \% change in the number of times Pakistan hosts diplomatic missions from Beijing in 2017. The first row indicates what this change is per a hypothetical percent increase aid from China to Pakistan, and the second row indicates what this change is per a hypothetical percent increase in loans (debt) from China to Pakistan. The marginal effect for aid on coverage in this case is \%0.76 while the marginal effect for debt on visits is \%0.05.

>>>>

---
title: "Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid and Debt, Xinhua and Diplomatic Visits"
author: ["Lucie Lu","Miles D. Williams"]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2: 
    toc: false
classoption: fleqn
fontfamily: mathpazo
fontsize: 12pt
bibliography: ref_xinhua.bib
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \setlength{\mathindent}{15pt}
  - \setlength\parindent{24pt}
  - \setlength{\parskip}{0.0pt plus 1.0pt}
  - \doublespacing
  - \newtheorem{theorem}{Theorem}[section]
  - \newtheorem{lemma}[theorem]{Lemma}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{corollary}[theorem]{Corollary}
  - \newenvironment{proof}[1][Proof]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries Proof}]}{\end{trivlist}}
  - \newenvironment{definition}[1][Definition]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{example}[1][Example]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{remark}[1][Remark]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{!t}
  - \usepackage{caption}\captionsetup[table]{skip=5pt}
  - \usepackage{dcolumn}
  - \usepackage{rotating}
  - \usepackage{caption}\DeclareCaptionLabelFormat{AppendixTables}{A.#2}

abstract: "Absent formal avenues of transparency, how do China’s foreign aid and debt allocation relate to public diplomacy and international recognition efforts? We bring a novel perspective to this issue by considering Chinese media coverage of developing countries intended for foreign audiences and China’s bilateral diplomatic activities. We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. We expect greater coverage of a developing country in a Chinese outward-facing media outlet and bilateral diplomatic activity to correlate with greater aid giving and debt allocation. To test this expectation, we leverage (1) AidData’s Chinese development finance datasets and (2) the meta-data of millions of Xinhua news articles in the Cline Center for Advanced Social Research. We find that China pays more media attention to aid recipients but engages more in diplomatic visits with debtors. Aid-recipient countries disproportionately get more media coverage on the state of their economy and developmental needs. Also, China is more likely to provide loans to diplomatically aligned countries. Our results dispel the myths of China as a ’rouge donor’ of desigining ’debt-trap diplomacy.’ We suggest that much like traditional donors, China’s foreign aid and debt allocations present a versatile foreign policy tool in international politics. "
---

@@

However, the goals of facilitating cooperation and winning international recognition as a rising power are difficult to accomplish in secret. This, therefore, raises a natural question. *How do China's foreign aid and debt allocation relate to its public diplomacy and international recognition efforts?*

We propose that despite the lack of formal reporting, Beijing can pursue international recognition and diplomacy with its overseas financing with the help of other complementary foreign policy activities. Specifically, we consider outward-facing state-sponsored news coverage and bilateral diplomatic missions from Beijing. Past research suggests that these two activities play special roles in complementing the foreign policy objectives that drive China's giving and lending.

@@

As we test these hypotheses, in our research design we delineate China's development financing according to the type of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.[^1] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

@@

We find that there are a few overlaps between the top 20 countries that receive aid and loans from Beijing at the same time. In other words, China target its aid and debt to different sets of developing countries. However, the neat distinction between China's use of aid and debt does not imply these tools serve separate goals, policy or commercial oriented ones, as conventional view holds. In particular, we find that *Xinhua* coverage is disproportionately concentrated among China's aid recipients, while no significant relationship exists between aid and diplomatic visits. Conversely, we find that debt is significantly linked to a greater frequency of bilateral diplomatic missions from Beijing, but no significant difference in *Xinhua* coverage. Together these findings reveal an intriguing pattern: Beijing likes to talk about its aid recipients but prioritizes diplomatic visits with its debtors.

On the one hand, aid recipient countries disproportionately get more media coverage from *Xinhua* via stories that report both on the state of their economy and their development needs. Notably, *Xinhua* does not just promote China's efforts to offer aid to those countries in need but also gives a rounded picture of those countries' economic conditions justifying their legitimate need for aid. On the other hand, China engages in more conventional forms of bilateral diplomacy via government-to-government diplomatic missions to cement closer bilateral ties with debtors.

These findings contradict with what conventional views hold that Beijing has tried to keep its giving secret. Instead, *Xinhua* devotes significant media attention and resources to Beijing's aid-recipients. It may help to promote Chinese roles and efforts in the developmental trajectories of these countries and improve Chinese image through visual branding of aid. Furthermore, China is not shy away from its ambition in building closer bilateral cooperation with its debt-recipients to align with their foreign policy goals. In addition to lending money, China engages in building and strengthening the friendships with these countries. Economic interests to motivate Beijing's debt allocations are suggestive to be secondary to its policy goals.

Meanwhile, the finding that loans corresponding with diplomatic missions has special relevance in light of Beijing's Belt and Road Initiative (BRI). A 2018 BRI foreign policy assessment report from Washington [@hurley_examining_2019] highlights that lending indicates Beijing's economic interests and strategic dominance---China tends to lend to vulnerable states with debt distress in exchange for unfair deals and strategic concessions. Brautigam [-@brautigam_critical_2020] calls the Chinese "debt-trap diplomacy" as a meme based on collective fear of the Chinese threat rather than sound empirical evidence. Our finding suggests that China is more likely to provide loans to diplomatically aligned countries. For instance, Russia, one of China's closest allies, receives the highest amount of lending on average among 142 countries but no aid at all between 2002 and 2017. In contrast, Iraq shows the opposite pattern: it receives the highest amount of aid from China during the period of our investigation, but it does not receive any loans. Concern about Chinese debt nonetheless raises a natural question related to the function of diplomacy in debtor countries. Does debt serve as a tool of diplomacy, or does Beijing wield diplomacy to cover a multitude of "debt-trap" sins? This study sheds some light on this pressing question, but further research is needed to fully answer it.

@@

Absent transparency in reporting, there are other tools at Beijing's disposal that may complement the diplomatic objectives behind its foreign aid. In this study, we explore the connections between China's bilateral development finance and other, *visible*, activities linked to diplomatic and legitimacy-seeking objectives. We focus in particular on two factors:  (1) coverage of developing countries in external-facing media (that is, media coverage intended for foreign, rather than domestic, audiences) and (2) elite level diplomatic visits from China to developing countries. We propose that countries receiving disproportionately China’s bilateral development finance support also receive more concentrated media coverage and diplomatic visits from Beijing.

We summarize these factors in greater detail in the following sections, but before considering them, it is necessary to first summarize the current landscape of China's overseas financing. In particular, it is important to draw a distinction between Beijing's ODA-like expenditures and its other official flows (OOF). The former we may call aid (or development assistance), and the latter we may call debt (or loans). As @dreher_banking_2022 note in their groundbreaking and comprehensive book on Chinese aid and loans, Beijing uses aid and debt to accomplish different goals, and it further does not finance them in equal proportion, both in terms of total spending and in terms of how it targets funds across developing countries. @dreher_banking_2022 argue that aid tends to be most associated with China's foreign policy objectives, while debt tends to correspond to its market interests. What do these differences imply for Beijing's bilateral diplomacy efforts and state-sponsored foreign media's reporting strategy?

On the one hand, since aid is supposed to have more direct links to China's foreign policy goals, it follows that China will complement its aid giving with visible acts of diplomacy and efforts to increase the salience of aid recipients in foreign audiences. On the other hand, while aid and debt may have distinct goals, such a tidy distinction may be too simplistic. Perhaps more importantly, some of the market oriented objectives that drive China's global debt might benefit from public displays of diplomacy. On this front, recent research suggests that Beijing uses its debt to ensure access to natural resources that China lacks domestically. As noted above, @dreherEtAl2021 find that Beijing targets more of its loans in resource rich developing countries, a strategy that @dreher_banking_2022 contend is meant to guarantee access to said resources. When Beijing's policy banks provide these loans, developing country governments collateralize their oil or rare minerals, all but assuring access to these resources for Beijing since many of these countries are unable to directly repay the loans [@dreher_banking_2022].

What this implies in the aggregate is that aid from China tends to be most strongly determined by both development and political factors, while debt is most strongly determined by market-based factors. For the two outcomes discussed in the following sections, this suggests that Chinese ODA-like expenditures more so than OOF should be strong predictors of other complementary and visible foreign policy activities. Meanwhile, given the more market-oriented (and by extension more controversial) application of debt, we should expect a much weaker association to exist between OOF and the factors summarized below. While visibility may be a virtue for aid, the opposite may be true for debt, if the two indeed serve separate goals. The next section summarizes these outcomes and their relevance as measures of foreign policy-related activities in the context of China's overseas financing.

@@

We have made the case that both foreign state-sponsored media and diplomatic visits can be used to sway foreign public opinion in favor of Beijing. Media coverage is one way that countries can generate attention and promote a positive spin on their activities to foreign audiences. As Nye [-@nye_public_2008-2] notes, daily communications that reach both domestic and foreign populations are a way to improve a country's soft power. Perhaps in recognition of this, Beijing has expanded its outward-facing state-sponsored media activities and news reporting over the past two decades. This has been done in part through the English version of *Xinhua*, China's top state-sponsored news agency. By covering developing countries that receive Chinese development financing, *Xinhua* editors can paint a positive picture of Beijing's financing and, in so doing, take steps to promote China's international image.

In addition to expanding the reach and scope of its externally facing news media, China engages in more conventional forms of bilateral diplomacy via government-to-government diplomatic missions. Efforts to document Beijing's public diplomacy have revealed an expansive set of activities in which Beijing engages. These range from the financial to the cultural, and from elite-to-elite meetings to informational exchanges. By sending regular diplomatic missions to developing countries, Beijing is able to strengthen its bilateral relationships with countries in the Global South helping to cement the close bilateral ties the offer of development assistance is, in part, indented to promote.

@@

 include_graphics(
   here('06_figures/coverages_combined.jpg'))

@@

 include_graphics(
   here('06_figures/visits_combined.jpg'))

@@

To put these estimates in more substantive terms, it will be helpful to show some computed marginal effects. Table \@ref(tab:mareff) shows the predicted change in each of the outcomes given a percent increase in either aid or debt in Pakistan in 2017. Cell entries in the first column denote the \% change in the number of times Pakistan appears in *Xinhua* in 2017 while entries in the second column denote the \% change in the number of times Pakistan hosts diplomatic missions from Beijing in 2017. The first row indicates what this change is per a hypothetical percent increase aid from China to Pakistan, and the second row indicates what this change is per a hypothetical percent increase in loans (debt) from China to Pakistan. The marginal effect for aid on coverage in this case is 0.76\% while the marginal effect for debt on visits is 0.05\%.
